---
title: "BayNorm_correlations"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(BiocManager)
library(bayNorm)
```

# BayNorm

### Background

Counts of gene transcripts can be observed within cells, but current methods have a low 'capture efficiency' (e.g. 10%) of true counts observed, leading to misleading count values that are lower and often zero/missing (dropout).

The capture efficiency $\beta_{j}$ varies across cells j, but we assume that gene-specific effects are small (i.e. transcripts from any gene i have roughly proportion $\beta_{j}$ of being captured in cell j). The capture efficiency also varies between experiments, 'batch-effects', where different batches of cells have different capture rates due to many factors (environment, slight differences in experiment, time etc).

### Method

Use a bayesian approach to estimate posterior distribution of true/original transcript counts $x_{i,j}^{0}$.

Prior: assume the true counts $x_{i,j}^{0}$ follow negative binomial distribution with parameters $\mu_i$, $\phi_i$ estimated using counts for gene i across all cells (or cells in the same batch: 'local')

Likelihood: assume that the observed count $x_{i,j}$ follows a binomial distribution with N = $x_{i,j}^{0}$ and p = $\beta_j$ (natural choice: each transcript has chance $\beta_j$ of being observed)

This gives the posterior as a shifted negative-binomial distribution.

### Application

Given the posterior distribution of the true counts $x_{i,j}^{0}$ for each gene i in each cells j there are several choices on how to proceed.

(1) MAP: 'maximum a posteriori probability', take the mode of the posterior as estimate of the true count $x_{i,j}^{0}$ for all i, j, producing a 'normalized' dataset of the same size as the original

(2) mean of posterior: as above, an estimate of the true count for each observed count

(3) sample from posterior: for each (i, j) sample S count values from the posterior e.g. S = 20. This produces a dataset of the same size as the observed counts BUT each entry is a list of S samples from the corresponding posterior. These can be considered as 'artifical cells', giving a dataset of 20 * the number of original cells to analyse.

# Data

Important to remove all missing values (NaN) otherwise the method will fail. Either drop genes/cells containing missing values, or preferably replace them by 0.

```{r}
# load dataset:
# genes with mean expression < 1 (and those that are all zero) have been removed
data_FIB = read.csv("C:/Users/willi/Documents/ProjectPaper/M5R/Interaction/Real-Data/Datasets/GSE151334_FIB_counts_thresh.csv", row.names=1)

# display
head(data_FIB, 5)
```

```{r}
# estimate beta (capture efficiency)
BETA_FIB <- bayNorm::BetaFun(Data=data_FIB, MeanBETA = 0.1)

# display
summary(BETA_FIB$BETA)
```

## MAP counts

```{r}
# MAP of posterior for each data  entry
bayNorm_2D_FIB <- bayNorm(
    Data = data_FIB,
    BETA_vec = NULL,
    mode_version= TRUE,
    mean_version = FALSE,
    verbose = TRUE,
    parallel = TRUE)
```

```{r}
# extract MAP counts
MAP_FIB <- data.frame(bayNorm_2D_FIB$Bay_out)

# display
head(MAP_FIB, 5)
```

```{r}
# store
write.csv(MAP_FIB, "C:/Users/willi/Documents/ProjectPaper/M5R/Interaction/Real-Data/Datasets/GSE151334_FIB_counts_MAP.csv")
```

```{r}
# first gene
par(mfrow=c(2,2))
# histogram of observed counts
hist(as.numeric(data_FIB[1,]), breaks=10, main="Observed Counts",
     xlab = "transcript count")
hist(as.numeric(data_FIB[1,]), breaks=100, main="Observed Counts",
     xlab = "transcript count")
# histogram of MAP counts
hist(as.numeric(MAP_FIB[1,]), breaks=10, main="MAP Count Estimates",
     xlab = "transcript count")
hist(as.numeric(MAP_FIB[1,]), breaks=100, main="MAP Count Estimates",
     xlab = "transcript count")
```

## Posterior Mean counts

```{r}
# Mean of posterior for each data entry
bayNorm_2D_FIB_MEAN <- bayNorm(
    Data = data_FIB,
    BETA_vec = NULL,
    mode_version= FALSE,
    mean_version = TRUE,
    verbose = TRUE,
    parallel = TRUE)
```

```{r}
# extract PM counts
MEAN_FIB <- data.frame(bayNorm_2D_FIB_MEAN$Bay_out)

# display
head(MEAN_FIB, 5)
```

```{r}
# store
write.csv(MEAN_FIB, "C:/Users/willi/Documents/ProjectPaper/M5R/Interaction/Real-Data/Datasets/GSE151334_FIB_counts_MEAN.csv")
```

```{r}
# first gene
par(mfrow=c(2,2))
# histogram of observed counts
hist(as.numeric(data_FIB[1,]), breaks=10, main="Observed Counts",
     xlab = "transcript count")
hist(as.numeric(data_FIB[1,]), breaks=100, main="Observed Counts",
     xlab = "transcript count")
# histogram of MAP counts
hist(as.numeric(MEAN_FIB[1,]), breaks=10, main="MEAN Count Estimates",
     xlab = "transcript count")
hist(as.numeric(MEAN_FIB[1,]), breaks=100, main="MEAN Count Estimates",
     xlab = "transcript count")
```

## Posterior Sample counts

```{r}
# S = 20 samples from posterior for each data  entry
bayNorm_3D_FIB <- bayNorm(
    Data=data_FIB,
    BETA_vec = NULL,
    mode_version=FALSE,
    mean_version = FALSE,
    S = 20,
    verbose =TRUE,
    parallel = TRUE)
```
```{r}
# extract counts
PS_FIB <- data.frame(bayNorm_3D_FIB$Bay_out)

# new sample sizes
print(dim(PS_FIB))

# display
head(PS_FIB, 5)
```
```{r}
# store
write.csv(PS_FIB, "C:/Users/willi/Documents/ProjectPaper/M5R/Interaction/Real-Data/Datasets/GSE151334_FIB_counts_PS.csv")
```

```{r}
# gene
par(mfrow=c(2,2))
j <- 1484
# histogram of observed counts
hist(as.numeric(data_FIB[j,]), breaks=10, main="Observed Counts",
     xlab = "transcript count")
# histogram of MAP counts
hist(as.numeric(MAP_FIB[j,]), breaks=100, main="MAP Count Estimates",
     xlab = "transcript count")
# histogram of posterior sample counts
hist(as.numeric(PS_FIB[j,]), breaks=100, main="Posterior Sample Counts",
     xlab = "transcript count")
hist(1)
```








