---
title: "CellCycleScoringCode"
output: html_document
---
```{r setup}
    knitr::opts_knit$set(root.dir = normalizePath("C:/Users/willi/Documents/ProjectPaper/M5R/Interaction/Latent-Experiments/R-Notebooks/")) 
```

```{r, echo=T}
library(anndataR)
library(reticulate)
library(Seurat)
library(Matrix)
library(dplyr)
library(cowplot)
library(openxlsx)
library(ggplot2)
library(edgeR)
library(scales)
library(reshape2)

#set colors
mycols<-c("#F39B7FB2","#E64B35B2", "#DC0000B2","#4DBBD5B2","steelblue2", "#3C5488B2","mediumturquoise","mediumaquamarine","forestgreen","#8491B4B2" )
```

# Cell Cycle Scoring

Following data pre-processing and analysis in "Part1.Human.cell.types.Rmd", including cell cycle scoring over all cells, use code from "Part3.Cell.cycle.markers.Rmd" to score cell cycle phases for each cell type.

```{r,echo=T}
counts<-readRDS("count.Seurat.obj.rds")
```

```{r,echo=T}
### Get biotypes
gtf<-read.delim(file = "GRC/GRCh38.id.name.biotype.tsv", sep='\t',  header=T)
gtf<-gtf[match(rownames(counts),gtf$external_gene_name), ] 
gtf<-data.frame(lapply(gtf, as.character),stringsAsFactors=FALSE)
```

#### Cell cycle scoring is performed on each cell type individually:
#### 1. Subset seurat object into three objects, one per cell type
#### 2. Assign a cycle phase to each cell
```{r,fig.width=5, fig.asp=.8, echo=T}
Idents(counts)<-"celltype"
counts.fibro<-subset(counts, idents = 'Fibroblasts')
counts.hek<-subset(counts, idents = 'HEK293T')
counts.mcf<-subset(counts, idents = 'MCF7')
```

```{r,message=FALSE, warning=FALSE}
cc.genes<-read.delim(file = "regev_lab_cell_cycle_genes.txt", sep="\t", header= F, stringsAsFactors = F)
s.genes<-cc.genes[1:43,]
g2m.genes<-cc.genes[44:97,]
```

```{r,message=FALSE, warning=FALSE}
counts.fibro <- CellCycleScoring(counts.fibro, s.features = s.genes, g2m.features = g2m.genes, set.ident = TRUE)
counts.mcf <-CellCycleScoring(counts.mcf, s.features = s.genes, g2m.features = g2m.genes, set.ident = TRUE)
counts.hek <-CellCycleScoring(counts.hek, s.features = s.genes, g2m.features = g2m.genes, set.ident = TRUE)
```

# Convert to AnnData

```{r}
counts_fibro_ad = as_AnnData(counts.fibro)
counts_mcf_ad = as_AnnData(counts.mcf)
counts_hek_ad = as_AnnData(counts.hek)
```

```{r}
write_h5ad(counts_fibro_ad, path = "anndata_FIB.h5ad")
write_h5ad(counts_mcf_ad, path = "anndata_MCF.h5ad")
write_h5ad(counts_hek_ad, path = "anndata_HEK.h5ad")
```

